---
author: "Dr. Raúl Trujillo Álvarez"
date: "`r format(Sys.time(), '%d %B, %Y')`"

output: 
   html_document:
    self-contained: true
---

## Libraries

-   `EPANET` is a computer program that performs extended period simulation of hydraulic and water quality behavior within pressurized pipe networks.EPANET Version 2.2 includes contributions from EPA and individuals outside the United States Government.

-   `epanet2toolkit` is an R package for simulating water networks using Epanet. The package provides functions from the Epanet programmer's toolkit as R functions so that basic or customized simulations can be carried out from R. The package uses Epanet version 2.2 from Open Water Analytics.

-   `epanetReader` is an R package for reading water network simulation data in Epanet's .inp and .rpt formats into R. Some basic summary information and plots are also provided.

```{r}
library(tidyverse)
library(here)
library(glue)
library(hyd4gpv)
library(epanet2toolkit)
library(epanetReader)

library(kableExtra)
```

## Define base parameters

-   Epanet Hydraulic model (.inp file): "exercise_01_Res2Res"
-   Valve brand: "AVK-ACMO"
-   Valve type: "PSEG"
-   Valve name in Hydraulic model: "RIKO"

```{r}
params <- list(
  file.hmodel = "exercise_01_Res2Res",
  valve.brand = "AVK-ACMO",
  valve.cyl   = "PSEG",
  valve.id    = "RIKO" # Name in the epanet .inp file
)
```

## Read basic about the control valve

```{r}
# Read Epanet model file; name in "params$file.hmodel"
file.hmodel <- read.inp(here::here("epanet",
                                   paste0(params$file.hmodel,".inp")))

# Read  from the epanet file the parameter of the control valve to evaluate.
# The name of the control valve is in "params$valve.id"
valve <- file.hmodel$Valves  |>  
  filter(ID == params$valve.id)

# Read Diameter of the valve.
params$valve.dn <- valve$Diameter[1]

# Read the nodes of the Control valve to analyze
params$valve.nodes <- c(valve[1,2], valve[1,3])

# Remove control.valve
rm(valve)

```

## Select the data kv curve parameters

```{r}
# Filter the control valves parameters table according to the brand, 
# cylinder type, and diameter..
params$valve.params <- hyd4gpv:::gpv_data |> 
  filter(brand == params$valve.brand & 
         cyl_name == params$valve.cyl) |> 
  filter(dn_min <= params$valve.dn &
         params$valve.dn <= dn_max)
```

## Generate the Zeta curve y generate the controls

-   Defines **controls** that modify the **TCV** (throttle control valve) based on the Loss Coefficient ($\zeta$)
-   The control changes the loss coefficient between $5\%$ and $100\%$ for each $5\%$ valve opening.

```{r}
# This part has the purpose of defining the controls that modify the 
# loss coefficient of the control valve (link TCV throttle control valve) 
# for the different opening positions of the valve (see file.hmodel$Times down).
zv4tcv <- tibble(position = seq(5, 100, by = 5)) |>
  mutate(zv = zv_function(position, 
                          params$valve.params$kv_b, 
                          params$valve.params$kv_d, 
                          params$valve.params$kv_e, 
                          params$valve.params$zvs)) |>
  select(position, zv) |>
  mutate(zv = round(zv,2)) |> arrange(zv) |>
  mutate(time = seq(1, length(zv), 1)) |>
  mutate(controls = glue('LINK {params$valve.id} {zv} AT TIME {time}; {position}% opening')) |>
  pull(controls)

# 1. Add the open position for the control valve.
#    This represent Without loses/valve at the control valve position
# 2. Change the information in the epanet file.
file.hmodel$Controls <- c(
  glue('LINK {params$valve.id} OPEN AT TIME 0; without Valve'), 
  zv4tcv
)

# Remove zv4tcv
rm(zv4tcv)

```

```{r}
file.hmodel$Controls |> 
  kbl() |> 
  kable_styling()
```

## Parameter for the report generation

```{r}
# Trust is good, but control is better (see controls definitions)
# length of Controls (zv4tcv)  must be the same of the duration Time:
file.hmodel$Times <- c(
  glue("Duration {length(file.hmodel$Controls)-1}:00"),
  "Hydraulic Timestep 1:00",
  "Quality Timestep 0:05",
  "Pattern Timestep 1:00",
  "Pattern Start 0:00",
  "Report Timestep 1:00",
  "Report Timestep 0:00",
  "Start ClockTime 12 am",
  "Statistic NONE"
)

file.hmodel$Report <- c(
  "Status  FULL",
  "Summary Yes",
  "Energy Yes",
  "Page 0",
  "Nodes All",
  "Links All",
  "Pressure Precision 4",
  "Velocity Precision 4",
  "Flow Precision 4",
  "Headloss Precision 4"
)

```

## Run modified hydraulic model

-   Write the hydraulic model object to a .inp file

-   Runs the complete EPANET simulation

-   Reads the Epanet .rpt file into R

```{r}
# Write an epanet.inp object to params$file.hmodel
write.inp(
  file.hmodel, 
  here::here("epanet", paste0(params$file.hmodel,".inp"))
)

# runs a complete EPANET simulation
ENepanet(
  here::here("epanet",paste0(params$file.hmodel,".inp")),
  here::here("epanet",paste0(params$file.hmodel,".rpt"))
)

# reads an Epanet .rpt file into R
file.report <- read.rpt(here::here("epanet","exercise_01_Res2Res.rpt")) 

```

## Reading the upstream ($p_1$) and downstream pressure ($p_2$) at the Control Valve.

```{r}
# Values for the Nodes of the Valve
nodes <- file.report$nodeResults  |> 
  filter(ID %in% params$valve.nodes) |> 
  select(ID, Pressure, position = Timestamp) |> 
  mutate(position = as.integer(str_extract(position, "^(\\d+)*")),
         Pressure = Pressure/10) |> 
  pivot_wider(names_from = ID, values_from = Pressure)  

names(nodes) <- c("position", "p1", "p2")

```

```{r}
nodes |> 
  kbl(caption = "Upstream and Downstream Pressure in bar") |> 
  kable_paper(bootstrap_options = "striped", full_width = F)
```

## Extracting Flow and velocity for the Control Valve

```{r}
# Values for the Valve (Link))
links <- as_tibble( file.report$linkResults) |>  
  filter(ID == params$valve.id) |> 
  select(
    position = Timestamp, 
    dh.valve = Headloss,       # in m
    flow = Flow,         # in lps
    velocity = Velocity  # in m/s
  ) |> 
  mutate(
    position = as.integer(str_extract(position, "^(\\d+)*")),
    flow = flow * 3.6, # convert Flow from lps to m3/h
    dh.valve = dh.valve/10 # Convert Dp from m to bar
  ) 

```

```{r}
links |> 
  kbl(caption = "Values of Flow (m3/h), Velocity (m/s) and Head losses (bar) for the Control Valve") |> 
  kable_paper(bootstrap_options = "striped", full_width = F)
```

## Calculation of the $k_v$ for every possition

```{r}
# join NODES with the LINKS
params$operat.values <- nodes |> 
  full_join(links, by ="position") |> 
  arrange(dh.valve)

# Extract the flow without valve.
params$system.maxflow <- params$operat.values$flow[1]

# Table with the Kv values 
params$operat.values <- params$operat.values |> 
  filter(dh.valve > 0) |> 
  mutate(
    position = seq(100, 5, by = -5),
    kv.valve = kv(p1, p2, flow),
    kv.valve.rel = kv.valve/max(kv.valve)*100
  )

rm(nodes, links)

```

```{r}
params$operat.values |> 
  kbl(caption = "Kv and Kv/Kvs values") |> 
  kable_paper(bootstrap_options = "striped", full_width = F) |> 
  footnote(general = "The Kv value expresses the amount of flow in a regulating valve at a given valve position with a pressure loss of 1 bar. The special situation with a fully open valve determines the Kvs value. ")
```
